Reproducibility Instructions

### 2021-01-09 R/README.Rmd
update md
edit section 4 & merge section 3 into section 2

---
title: "Road Segment Prioritization for Bicycle Infrastructure"
author: "Hussein Mahfouz"
date: "18/8/2020"
output: github_document
fontsize: 10pt
---

@@

There are a couple of files that cannot be synced to github due to their size. These files are neseccary for the scripts to run. Below are links to where you can download them, and instructions on where to place them in the repo file structure

@@

The script then filters all flow data where both the Origin MSOA **AND** the destination MSOA are in the chosen city

If you wish to run the analysis on London, then make sure you have a computer that is up to the task. I didn't :(

@@

This script is used to estimate where additional cycling demand will come from. Let's say that the target for Manchester is a 10% increase in cyling mode share, how many additional cyclists do we assign to each OD pair to reach that target mode share

1. Predict Probability of Cycling Between Each OD Pair Based On Geography
      1. A glm is used to predict probability of cycling based on distance and slope

2. Accounting for Existing Mode Share
      1. Look at performance of each OD pair and assign additional cyclists accordingly. OD pairs that have a            low cycling mode share are allocated more cyclists than OD pairs that already have a high cycling mode          share. This is because OD pairs with low cycling mode share have more potential (latent demand) than OD          pairs with high cycling mode share

3. Scaling Results To Match Mode Share Target
      1. Specify target mode share increase (default is 10% but this is unreasonable for a city like Cambridge           that already has a cycling mode share of 40%)
      2. Scale potential cycling demand up or down so that it matches the target % increase

@@

* Compare the distance distribution of existing cycling mode share and potential cycling mode share:

<p float="left">
  <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling.png" width="250" />
  <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png" width="250" /> 
  <img src="../data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" width="250" />
</p>

* Vizualize Existing and Potential Cycling Flow as Desire Lines

![desire lines](../data/Manchester/Plots/desire_facet_cycling.png)

* Examine where potential cycling demand is assigned.

The methodology in ___3.0_potential_demand.R_ insures that OD pairs that have a low cycling mode share are allocated more cyclists than OD pairs that already have a high cycling mode share. In the figure below, the x axis is a ratio of the cycling mode share of the OD pair to its expected cycling mode share. The expected cycling mode share is obtained from a glm where distance, sqrt(distance), and slope are used as predictors. Looking at the resulting cycling mode share, we see that OD pairs between 2-8km have the highest mode share (consistent with bell-shaped distribution of cycling vs distance), and that mode share increase is highest for OD pairs that have lower than expected cycling mode shares.

<p float="center">
  <img src="../data/Manchester/Plots/mode_share_increase_vs_performance_smooth_Manchester.png" width="400" />
</p>

@@

These weighting profiles are used in script ___4.0_aggregating_flows.R_

@@

Here we analyze the street network configuration of the city by comparing the unweighted shortest paths to the weighted shortest paths (check methodology for explanation of weighted shortest paths). The aggregated flow shows us which road types are used, and it is clear that cycleways are not utilized unless the road network is weighted to create a hierarchy of road type preference.

![Unweighted Routing](../data/Manchester/Plots/flows_facet_unweighted.png)

![Weighted Routing](../data/Manchester/Plots/flows_facet_weighted.png)

@@

Using the potential cycling demand between OD pairs, we are able to define communities in the network. The nodes are the population-weighted MSOA centroids (location obtained from the [pct](https://itsleeds.github.io/pct/) package) and the links between them are weighted by the potential cycling demand between them. The Louvian algorithm is used to assign each MSOA centroid to a community, and then each road segment on the network is assigned to the same community as the MSOA centroid closest to it. The results for Manchester are shown below

![Community Detection](../data/Manchester/Plots/communities_alternative_Manchester.png)

@@

Results:

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png" width="350" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column.png" width="350" /> 
</p>

The results show the priority of each road segment (Roads are grouped into 100km groups for vizualization purposes)

<p float="center">
  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_priority_all_FLOW.png" width="450" />
</p>

@@

Results:

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png" width="350" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column.png" width="350" /> 
</p>

Priority of each road segment, and utilization of different OSM road types:

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png" width="300" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png" width="300" /> 
</p>

##### Comparing Connectivity of Algorithm 2 and 3

We check the number of connected components and the size of the Largest Connected Component as road segments are added to the solution (the components are the road segments). The initial number of components depends on the existing bicycle network of the city. For Manchester, we can see that the existing bicycle network has over 120 disconnected components (Remember we are only looking at segregated bicycle infrastructure, not painted bicycle lanes).

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png" width="350" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png" width="350" /> 
</p>

The algorithms seem to provide comparable connectivity gains.

>>>>

---
title: "Reproducibility Instructions"
#author: "Hussein Mahfouz"
date: "18/8/2020"
output: github_document
fontsize: 10pt
---

@@

There is one file that cannot be synced to github due to its size. This file is neseccary for the scripts to run:

@@

The script then filters all flow data where both the Origin MSOA **AND** the destination MSOA are in the chosen city. If you wish to run the analysis on London, then make sure you have a computer that is up to the task.

@@

This script is used to estimate where additional cycling demand will come from. It uses an uptake function from the Propensity to Cycle Tool ([pct](https://itsleeds.github.io/pct/)).

@@

<!-- * Compare the distance distribution of existing cycling mode share and potential cycling mode share: -->

<!-- <p float="left"> -->
<!--   <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling.png" width="250" /> -->
<!--   <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png" width="250" />  -->
<!--   <img src="../data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" width="250" /> -->
<!-- </p> -->

<!-- * Vizualize Existing and Potential Cycling Flow as Desire Lines -->

<!-- ![desire lines](../data/Manchester/Plots/desire_facet_cycling.png) -->

<!-- * Examine where potential cycling demand is assigned.  -->

<!-- In the figure below, the x axis is a ratio of the cycling mode share of the OD pair to its expected cycling mode share. Looking at the resulting cycling mode share, we see that OD pairs between 2-8km have the highest mode share (consistent with bell-shaped distribution of cycling vs distance), and that mode share increase is highest for OD pairs that have lower than expected cycling mode shares. -->

<!-- <p float="center"> -->
<!--   <img src="../data/Manchester/Plots/mode_share_increase_vs_performance_smooth_Manchester.png" width="400" /> -->
<!-- </p> -->

@@

These weighting profiles are used in script ___2.0_distance_and_elevation.R_ AND ___4.0_aggregating_flows.R_

@@

Here we analyze the street network configuration of the city by comparing the unweighted shortest paths to the weighted shortest paths (check methodology for explanation of weighted shortest paths). 
<!-- The aggregated flow shows us which road types are used, and it is clear that cycleways are not utilized unless the road network is weighted to create a hierarchy of road type preference.  -->

<!-- ![Unweighted Routing](../data/Manchester/Plots/flows_facet_unweighted.png) -->

<!-- ![Weighted Routing](../data/Manchester/Plots/flows_facet_weighted.png) -->

@@

Using the potential cycling demand between OD pairs, we are able to define communities in the network. The nodes are the population-weighted MSOA centroids (location obtained from the [pct](https://itsleeds.github.io/pct/) package) and the links between them are weighted by the potential cycling demand between them. The Louvian algorithm is used to assign each MSOA centroid to a community, and then each road segment on the network is assigned to the same community as the MSOA centroid closest to it.

<!-- The results for Manchester are shown below  -->

<!-- ![Community Detection](../data/Manchester/Plots/communities_alternative_Manchester.png) -->

@@

<!-- Results:  -->

<!-- <p float="left"> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png" width="350" /> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column.png" width="350" />  -->
<!-- </p> -->

<!-- The results show the priority of each road segment (Roads are grouped into 100km groups for vizualization purposes) -->

<!-- <p float="center"> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_priority_all_FLOW.png" width="450" /> -->
<!-- </p> -->

@@

<!-- Results:  -->

<!-- <p float="left"> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png" width="350" /> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column.png" width="350" />  -->
<!-- </p> -->

<!-- Priority of each road segment, and utilization of different OSM road types: -->

<!-- <p float="left"> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png" width="300" /> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png" width="300" />  -->
<!-- </p> -->

<!-- ##### Comparing Connectivity of Algorithm 2 and 3 -->

<!-- We check the number of connected components and the size of the Largest Connected Component as road segments are added to the solution (the components are the road segments). The initial number of components depends on the existing bicycle network of the city. For Manchester, we can see that the existing bicycle network has over 120 disconnected components (Remember we are only looking at segregated bicycle infrastructure, not painted bicycle lanes). -->

<!-- <p float="left"> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png" width="350" /> -->
<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png" width="350" />  -->
<!-- </p> -->

<!-- The algorithms seem to provide comparable connectivity gains.  -->
